---
title: "NOAA Rainfall"
author: "Joy-El Talbot"
date: "6/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# TODO change path to your working directory
WORKINGDIR <- "C:/Users/joyel/Documents/Roux/ALY6010/2021Sb_ALY6010/"

# R-markdown/Knitr alternative to setwd()
# MUST be run in this first chunk called "setup"
knitr::opts_knit$set(root.dir = WORKINGDIR) # sets working dir when Knitting

# TODO if coding live as you build this document copy & paste the setwd() command into your console.
setwd(WORKINGDIR) # sets working dir for R studio instance IF copied directly into console

# suppress scientific notation
options(scipen = 99)

# suppress loading strings as factors
options(stringsAsFactors = FALSE)
```

## Purpose
Quickly explore a new data set.

### Table of Contents
  
  1. Purpose
  2. Project Setup
  3. TODO add more "chapters" (use double-hash to label in text ##)
  
### Appendices
  
  1. Output Tables
  2. Output ggplot Graphs
  3. Output base-R Graphs


## Project Setup

### Directory Setup
```{r createProjectDirectory, eval=FALSE}
# TODO set eval=TRUE in line above to RUN this code
# TODO update as you desire for your directory
# TODO manually move data sources into the data/ folder
dir.create("data/")
dir.create("src/")
dir.create("analysis/")
dir.create("results/")
```

### Load Libaries
```{r libraries}
# TODO uncomment those you use in your subsequent code
library(ggplot2)
library(tidyr)
library(dplyr)
library(purrr) # part of tidyverse to get access to the map() function
```

### Define Constants
```{r constants}
# input source
DATADIR <- "data/"

# build results directory
# TODO name your analysis with the ANALYSISNAME constant
ANALYSISNAME <- "noaa_rainfall"  # NO spaces!

# TODO choose date vs. datetime label for your output folder
#DATESTAMP <- format(Sys.time(), "%Y-%m-%d")
DATESTAMP <- format(Sys.time(), "%Y-%m-%d-%H%M%S") # includes hours, minutes, and seconds if running multiple times in a day
RESULTSDIR <- paste0("results/",DATESTAMP,"-",ANALYSISNAME,"-results/")
dir.create(RESULTSDIR)
GRAPHDIR <- paste0(RESULTSDIR, "graphs/")
dir.create(GRAPHDIR)
TABLEDIR <- paste0(RESULTSDIR, "tables/")
dir.create(TABLEDIR)

# define default plot size
PLOT_WIDTH <- 5
PLOT_HEIGHT <- 5
PLOT_UNITS <- "in"
PLOT_RES <- 96 # required for png() device when units aren't "px"
```

### Load data
```{r loadData}
# TODO load your particular data (an example is below)
#checkouts <- read.csv(paste0(DATADIR,"sampled_dailyCheckouts.csv"),
#                      stringsAsFactors = FALSE)
rain <- utils::read.csv(paste0(DATADIR, "multicity_rain_noaa.csv"),
                        stringsAsFactors = FALSE)
```

## Analysis

### Initial Viewing
```{r exploration}
head(rain)
rain[0:10,]
rain[2000:2010,]
summary(rain)
unique(rain$STATION_NAME)
```

### Data Cleanup

  1. Convert 999.99 HPCP values to NA
  2. Summarize to daily rainfall

```{r datacleanup}
clean_rain <- rain

clean_rain$HPCP[clean_rain$HPCP == 999.99] <- NA

summary(clean_rain)

clean_rain <- clean_rain[!is.na(clean_rain$HPCP), ]

summary(clean_rain)

clean_rain$day <- as.character(map(strsplit(clean_rain$DATE, " "), 1))

rain_summary <- clean_rain %>%
  group_by(day) %>%
  dplyr::summarise(dailyRain = sum(HPCP))

rain_summary$year <- substr(rain_summary$day, 0, 4)
```

```{r plotByyear}
df_plot <- rain_summary %>%
  group_by(year) %>%
  dplyr::summarise(averageDailyRain = mean(dailyRain)) %>%
  arrange(averageDailyRain)

(ggplot(df_plot)
 + aes(x = as.character(year), y = averageDailyRain)
 + geom_bar(stat="identity")
 + ggtitle("a plot")
 
 )
# TODO order by ascending ave daily rain
# TODO Does order of piping affect output
# TODO coord_cartesian vs xlim zoom
# TODO boxplot with jitter overlay
# TODO scatter for rain data (how to subset by year)
```


TODO add rounds of "Chapters" ##, "Sub-chapters" ###, "Sub-sub-chapters" ####, etc AND code blocks to build out your analysis.















## Appendix - Examples for Outputting code
TODO Feel free to remove if desired

### Output Tables
```{r outputTables, eval=FALSE}
byYear <- checkouts %>%
  group_by(CheckoutYear) %>%
  summarise(Count = n())
print(byYear)

write.csv(byYear,
          file = paste0(TABLEDIR,"Checkouts_by_year.csv"),
          row.names = FALSE) # to suppress row.names of 1, 2, 3...
```

  + `write.csv` ensures formatting of output matches CSV file standards
  + `write.table` gives more flexiblity but usually not needed
  + Other commands for writing JSON, XML, and other formats can be found with a quick search (and usually a library() package).


### Output Graph Images
#### With ggplot2
```{r outputGraphs_ggplot, eval=FALSE}
# wrap in parentheses so the "+" signs are a the left, making it easier to comment out lines
(ggplot(checkouts)
+ aes(x = factor(CheckoutYear))  # adding factor() so these names become x-axis label names
+ geom_bar()
+ ggtitle("Patrons checkout fewer books in 2020")
+ labs(subtitle = "Seattle Public Library - Physical Book Checkouts")
+ scale_y_continuous("Physical Books Checked Out\n(2% sample)")
#+ scale_x_discrete("") # no label
+ scale_x_discrete("Year")
)

# saves the last image you plotted to file by default
# OTHERWISE you can pass a ggplot object to the plot= argument
ggsave(filename = paste0(GRAPHDIR,"ggplot_CheckoutsByYear.png"),
       width = PLOT_WIDTH*0.9, # add multipliers to make image a little bigger or smaller as needed (if needed)
       height = PLOT_HEIGHT*1.1,
       units = PLOT_UNITS)
```

  + Graph first, then save after (similar to table output workflow)
  + `ggsave()` will write the last plot to file
  + Setting constants for plot size makes it easy to keep images of consistent size
  + Nudge size larger or smaller with a size multiplier to the PLOT_WIDTH or PLOT_HEIGHT constants


#### With base R plots
```{r outputGraphs_basePlot, eval=FALSE}
# open output file
png(filename = paste0(GRAPHDIR,"basePlot_CheckoutsByYear.png"),
    width = PLOT_WIDTH,
    height = PLOT_HEIGHT,
    units = PLOT_UNITS,
    res = PLOT_RES)

# create plot
hist(x = checkouts$CheckoutYear,
     main = "Patrons checkout fewer books in 2020",
     sub = "Seattle Public Library - Physical Book Checkouts",
     xlab = "Year",
     ylab = "Physical Books Checked Out\n(2% sample)")
      
# close the output device
dev.off()
```

  + Sandwich file-saving code around the graph-building code
  + If your graph building errors out, you may leave "devices" open 
    + Avoid by testing graph code first (without opening a device)
    + Fix by repeatedly run dev.off() in the console
