---
title: "NOAA Rainfall - Continued Analysis"
author: "Joy-El Talbot"
date: "6/3/2021"
output: html_document
---

updated: 6/6/2021 by Joy-El

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# TODO change path to your working directory
WORKINGDIR <- "C:/Users/joyel/Documents/Roux/ALY6010/2021Sb_ALY6010/"

# R-markdown/Knitr alternative to setwd()
# MUST be run in this first chunk called "setup"
knitr::opts_knit$set(root.dir = WORKINGDIR) # sets working dir when Knitting

# TODO if coding live as you build this document copy & paste the setwd() command into your console.
setwd(WORKINGDIR) # sets working dir for R studio instance IF copied directly into console

# suppress scientific notation
options(scipen = 99)

# suppress loading strings as factors
options(stringsAsFactors = FALSE)
```

## Purpose
Quickly explore a new data set.

## Assumptions

  1. HPCP *hourly rain data is in units of inches* having already been converted from hundredths of inches before we downloaded it.
  2. HPCP 999.99 values represent missing data and will be *excluded* from the analysis.

### Table of Contents
  
  1. Purpose
  2. Project Setup
  3. Analysis
  
### Appendices
  
  1. Output Tables
  2. Output ggplot Graphs
  3. Output base-R Graphs


## Project Setup

### Directory Setup
```{r createProjectDirectory, eval=FALSE}
# TODO set eval=TRUE in line above to RUN this code
# TODO update as you desire for your directory
# TODO manually move data sources into the data/ folder
dir.create("data/")
dir.create("src/")
dir.create("analysis/")
dir.create("results/")
```

### Load Libaries
```{r libraries}
library(ggplot2)
library(tidyr)
library(dplyr)
library(purrr) # part of tidyverse to get access to the map() function
```

### Define Constants
```{r constants}
# input source
DATADIR <- "data/"

# build results directory
ANALYSISNAME <- "noaa_rainfall"  # NO spaces!

DATESTAMP <- format(Sys.time(), "%Y-%m-%d-%H%M%S") # includes hours, minutes, and seconds if running multiple times in a day
RESULTSDIR <- paste0("results/",DATESTAMP,"-",ANALYSISNAME,"-results/")
dir.create(RESULTSDIR)
GRAPHDIR <- paste0(RESULTSDIR, "graphs/")
dir.create(GRAPHDIR)
TABLEDIR <- paste0(RESULTSDIR, "tables/")
dir.create(TABLEDIR)

# define default plot size
PLOT_WIDTH <- 5
PLOT_HEIGHT <- 5
PLOT_UNITS <- "in"
PLOT_RES <- 96 # required for png() device when units aren't "px"
```

### Load data
```{r loadData}
rain <- utils::read.csv(paste0(DATADIR, "multicity_rain_noaa.csv"),
                        stringsAsFactors = FALSE)
```

## Analysis

### Initial Viewing
```{r exploration}
head(rain)
rain[0:10,]
rain[2000:2010,]
summary(rain)
unique(rain$STATION_NAME)
```

### Data Cleanup

  1. Convert 999.99 HPCP values to NA
  2. Summarize to daily rainfall
  3. Create day of year field in summary data

```{r datacleanup_step1}
# Step 1 - Convert 999.99 HPCP values to NA
clean_rain <- rain

clean_rain$HPCP[clean_rain$HPCP == 999.99] <- NA

summary(clean_rain)

clean_rain <- clean_rain[!is.na(clean_rain$HPCP), ]

summary(clean_rain)
```

#### Some notes about the HPCP column

According to the [data documentation from NOAA](https://www1.ncdc.noaa.gov/pub/data/cdo/documentation/PRECIP_HLY_documentation.pdf), the HPCP column is originally gathered as hundreths of inches of rainfall. The 99999 value is used to represent missing data values. 

Given the format of the 99999 value in our imported data as 999.99, we ASSUME that the data has been converted to inches by dividing all HPCP values by 100.


```{r datacleanup_step1b}
# Step 1b - determine units of HPCP and name accordingly
clean_rain <- rename(clean_rain, inchesRain = HPCP)
head(clean_rain)
```

```{r datacleanup_step2}
# Step 2 - Summarize to daily rainfall

# Step 2a - get date string from timestamp
clean_rain$day <- as.character(map(strsplit(clean_rain$DATE, " "), 1))

# Step 2b - group by date string and sum inchesRain to get daily rainfall
rain_summary <- clean_rain %>%
  group_by(day) %>%
  dplyr::summarise(dailyRainInches = sum(inchesRain))

rain_summary$year <- substr(rain_summary$day, 0, 4)
```

```{r datacleanup_step3}
# create field for day of year number
rain_summary$date <- as.Date(rain_summary$day, format = "%Y%m%d")

# strftime() solution from stackoverflow, hat-tip @Gavin Simpson; https://stackoverflow.com/questions/7958298/how-do-you-convert-posix-date-to-day-of-year
rain_summary$dayOfYear <- strftime(rain_summary$date, format = "%j") %>%
                            as.numeric()
head(rain_summary)
```


### Explore Data via Daily Rain field

#### Plot average daily rainfall by year
```{r plotByyear}
df_plot <- rain_summary %>%
  group_by(year) %>%
  dplyr::summarise(averageDailyRain = mean(dailyRainInches)) %>%
  arrange(desc(averageDailyRain))

head(df_plot)
write.csv(df_plot,
          file = paste0(TABLEDIR,"aveDailyRainbyYear.csv"),
          row.names = FALSE)

(ggplot(df_plot)
  # reorder() orders first field according to order of second field
  #   defaults to ascending order on second field
  #     override with desc() function around second field
 #+ aes(x = reorder(as.character(year), desc(averageDailyRain)), y = averageDailyRain)
 + aes(x = as.character(year), y = averageDailyRain)
 + geom_bar(stat = "identity")
 + ggtitle(label = "Daily rain averages remain consistent each year.",
           subtitle = "Average Daily Rain fall per year")
 + scale_x_discrete("Year")
 + scale_y_continuous("Average Daily Rainfall (Inches)")
 )

ggsave(filename = paste0(GRAPHDIR, "aveDailyRainbyYear.png"),
       width = PLOT_WIDTH,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)
```

#### Plot daily rainfall across each year
Goal: view line graph of daily rainfall for each year
  x-axis = days of year
  y-axis = total rain in inches
  group = color-code by year, maybe facet

Recommended reading - [examples of boxplots & jitterplots](https://cmdlinetips.com/2019/02/how-to-make-grouped-boxplots-with-ggplot2/)

```{r plotDailyRainTimeseries}
df_plot <- rain_summary
df_plot$unifiedDate <- as.Date(strptime(df_plot$dayOfYear, format = "%j"))

(ggplot(df_plot)
 + aes(x = unifiedDate,
       y = dailyRainInches,
       group = year)
 + geom_line(aes(color = year))
 + facet_grid(year ~ .)
 + theme(legend.position = "none")
 + ggtitle(label = "Daily Rain fall highlights wetter and drier periods.",
           subtitle = "Daily Rain fall per year")
 + scale_x_date("Day of Year", 
                date_labels = "%b",
                date_breaks = "1 month")
 + scale_y_continuous("Daily Rainfall (Inches)")
 )

# Finding it hard to identify patterns in the time series (too much noise, eg.
# too much variability). Instead going to try plotting as boxplots to condense 
# down the variability and facilitate comparisons.

df_plot$month <- strftime(df_plot$unifiedDate, format = "%b")
df_plot$month <- factor(df_plot$month,
                        levels = c("Jan", "Feb", "Mar", "Apr",
                                   "May", "Jun", "Jul", "Aug",
                                   "Sep", "Oct", "Nov", "Dec"),
                        ordered = TRUE)
head(df_plot)
(ggplot(df_plot)
  + aes(x = month,
        y = dailyRainInches,
        fill = factor(year))
  + geom_boxplot()
  + scale_x_discrete("Month")
  + scale_y_continuous("Daily Rain (Inches)")
  + coord_cartesian(ylim = c(0,3))
  + ggtitle(label = "Daily Rain fall highlights wetter & drier periods.",
            subtitle = paste0("Daily Rain fall by month & year\nExlcudes ",
                              sum(rain_summary$dailyRainInches > 3),
                              " days with > 3 inches of rain"))
  + scale_fill_discrete("Year")
  + theme(legend.position = "bottom")
  )
ggsave(filename = paste0(GRAPHDIR, "dailyRainbyMonthYear_boxplot.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

```

#### Compare plot types - histogram, bar graph, and boxplot
Purpose: show the same (or as close the same as possible) data with all three plot types.
```{r compareHistBarBox}
# Histogram shows frequency of each value for a continuous variable
#  in this case, dailyRainInches from the rain_summary table will be our continuous variable
(ggplot(rain_summary)
 + aes(x = dailyRainInches)
 + geom_histogram(bins = 20) # I chose 20 pretty arbitrarily
 + scale_x_continuous("Daily Rain (Inches)")
 + scale_y_continuous("Number of Days")
 + ggtitle("Most days experienced little rainfall",
           subtitle = "Distribution of Rain by Day (2005-2010)")
 )

ggsave(filename = paste0(GRAPHDIR, "dailyRain_hist.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

# let's exclude days with NO rain
no_rain_days <- sum(rain_summary$dailyRainInches == 0)
(ggplot(rain_summary[rain_summary$dailyRainInches > 0, ])
 + aes(x = dailyRainInches)
 + geom_histogram(bins = 20) # I chose 20 pretty arbitrarily
 + scale_x_continuous("Daily Rain (Inches)")
 + scale_y_continuous("Number of Days")
 + ggtitle("Most days experienced little rainfall",
           subtitle = paste0("Distribution of Rain by Day (2005-2010); excludes ",
                             no_rain_days,
                             " days without any recorded rain."))
 )
ggsave(filename = paste0(GRAPHDIR, "dailyRain_excludeNoRain_hist.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

# Bar graph shows a summary (typically count or average) of a continuous variable 
# across multiple categorical values. 
#   in this case, dailyRainInches will be our continuous variable, year will be our
#   categorical, and we will show the average dailyRain for each year.

# geom_bar defaults to counting each item provided, to get other types of summaries
# we need to do these ourselves.
df_plot <- rain_summary %>%
  group_by(year) %>%
  summarise(aveRain = mean(dailyRainInches))

(ggplot(df_plot)
  + aes(x = year, y = aveRain)
  + geom_bar(stat = "identity") # sending stat "identity" to tell ggplot to use the raw values
  + scale_x_discrete("Year")
  + scale_y_continuous("Average Daily Rain (Inches)")
  + ggtitle("Average daily rain has changed little over the years",
            subtitle = "Average Daily Rain by Year")
  )

ggsave(filename = paste0(GRAPHDIR, "dailyRainByYear_bar.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

# Boxplots summarize the distribution of a continuous variable (specifically 
# median, IQR, and 1.5 IQR) across the entire dataset (one boxplot) or for multiple
# categorical values.
#   in this case, dailyRainInches will be our continuous variable, year will be our
#   categorical

(ggplot(rain_summary)
  + aes(x = year, y = dailyRainInches)
  + geom_boxplot()
  + scale_x_discrete("Year")
  + scale_y_continuous("Daily Rain (Inches)")
  + ggtitle("Daily Rain tends to fall below 2.5 inches / day",
            subtitle = "Daily Rain by Year (2005 - 2010)")
  )
ggsave(filename = paste0(GRAPHDIR, "dailyRainByYear_box.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

# let's exclude days with NO rain
no_rain_days <- sum(rain_summary$dailyRainInches == 0)
(ggplot(rain_summary[rain_summary$dailyRainInches > 0, ])
  + aes(x = year, y = dailyRainInches)
  + geom_boxplot()
  + scale_x_discrete("Year")
  + scale_y_continuous("Daily Rain (Inches)")
  + ggtitle("Daily Rain tends to fall below 2.5 inches / day",
            subtitle = paste0("Daily Rain by Year (2005-2010); excludes ",
                             no_rain_days,
                             " days without any recorded rain."))
 )
ggsave(filename = paste0(GRAPHDIR, "dailyRainByYear_excludeNoRain_box.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)

# Let's go back to the first box plot, but zoom in on the boxes
(ggplot(rain_summary)
  + aes(x = year, y = dailyRainInches)
  + geom_boxplot()
  + scale_x_discrete("Year")
  + scale_y_continuous("Daily Rain (Inches)")
  + ggtitle("Daily Rain tends to fall below 2.5 inches / day",
            subtitle = "Daily Rain by Year (2005 - 2010); does not show points above 2.5 inches")
  + coord_cartesian(ylim = c(0, 2.5))
  )
ggsave(filename = paste0(GRAPHDIR, "dailyRainByYear_zoomed_box.png"),
       width = PLOT_WIDTH*2,
       height = PLOT_HEIGHT,
       units = PLOT_UNITS)
```


#### Plot March 2010 data vs other years
Scatter plots to compare March rainfall across years.
# TODO make this plot!


TODO add rounds of "Chapters" ##, "Sub-chapters" ###, "Sub-sub-chapters" ####, etc AND code blocks to build out your analysis.















## Appendix - Examples for Outputting code
TODO Feel free to remove if desired

### Output Tables
```{r outputTables, eval=FALSE}
byYear <- checkouts %>%
  group_by(CheckoutYear) %>%
  summarise(Count = n())
print(byYear)

write.csv(byYear,
          file = paste0(TABLEDIR,"Checkouts_by_year.csv"),
          row.names = FALSE) # to suppress row.names of 1, 2, 3...
```

  + `write.csv` ensures formatting of output matches CSV file standards
  + `write.table` gives more flexiblity but usually not needed
  + Other commands for writing JSON, XML, and other formats can be found with a quick search (and usually a library() package).


### Output Graph Images
#### With ggplot2
```{r outputGraphs_ggplot, eval=FALSE}
# wrap in parentheses so the "+" signs are a the left, making it easier to comment out lines
(ggplot(checkouts)
+ aes(x = factor(CheckoutYear))  # adding factor() so these names become x-axis label names
+ geom_bar()
+ ggtitle("Patrons checkout fewer books in 2020")
+ labs(subtitle = "Seattle Public Library - Physical Book Checkouts")
+ scale_y_continuous("Physical Books Checked Out\n(2% sample)")
#+ scale_x_discrete("") # no label
+ scale_x_discrete("Year")
)

# saves the last image you plotted to file by default
# OTHERWISE you can pass a ggplot object to the plot= argument
ggsave(filename = paste0(GRAPHDIR,"ggplot_CheckoutsByYear.png"),
       width = PLOT_WIDTH*0.9, # add multipliers to make image a little bigger or smaller as needed (if needed)
       height = PLOT_HEIGHT*1.1,
       units = PLOT_UNITS)
```

  + Graph first, then save after (similar to table output workflow)
  + `ggsave()` will write the last plot to file
  + Setting constants for plot size makes it easy to keep images of consistent size
  + Nudge size larger or smaller with a size multiplier to the PLOT_WIDTH or PLOT_HEIGHT constants


#### With base R plots
```{r outputGraphs_basePlot, eval=FALSE}
# open output file
png(filename = paste0(GRAPHDIR,"basePlot_CheckoutsByYear.png"),
    width = PLOT_WIDTH,
    height = PLOT_HEIGHT,
    units = PLOT_UNITS,
    res = PLOT_RES)

# create plot
hist(x = checkouts$CheckoutYear,
     main = "Patrons checkout fewer books in 2020",
     sub = "Seattle Public Library - Physical Book Checkouts",
     xlab = "Year",
     ylab = "Physical Books Checked Out\n(2% sample)")
      
# close the output device
dev.off()
```

  + Sandwich file-saving code around the graph-building code
  + If your graph building errors out, you may leave "devices" open 
    + Avoid by testing graph code first (without opening a device)
    + Fix by repeatedly run dev.off() in the console
