---
title: "Teacher File - R Practice Week 3"
author: "Joy-El Talbot"
date: "6/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# TODO change path to your working directory
WORKINGDIR <- "C:/Users/joyel/Documents/Roux/Data/seattle_library/"

# R-markdown/Knitr alternative to setwd()
# MUST be run in this first chunk called "setup"
knitr::opts_knit$set(root.dir = WORKINGDIR) # sets working dir when Knitting

# TODO if coding live as you build this document copy & paste the setwd() command into your console.
setwd(WORKINGDIR) # sets working dir for R studio instance IF copied directly into console

# suppress scientific notation
options(scipen = 99)

# suppress loading strings as factors
options(stringsAsFactors = FALSE)
```

## Purpose
Prepare some population-level statistics for prompts in the Module 3 R Practice.

### Table of Contents
  
  1. Purpose
  2. Project Setup
  3. Population Statistics

## Project Setup

### Directory Setup
```{r createProjectDirectory, eval=TRUE}
dir.create("results/")
```

### Load Libaries
```{r libraries}
library(dplyr)
library(purrr)
```

### Define Constants
```{r constants}
# input source
DATADIR <- "data/"

# build results directory
ANALYSISNAME <- "ALY6010_Week3_RPractice"  # NO spaces!

DATESTAMP <- format(Sys.time(), "%Y-%m-%d")
RESULTSDIR <- paste0("results/",DATESTAMP,"-",ANALYSISNAME,"-results/")

# define default plot size
PLOT_WIDTH <- 5
PLOT_HEIGHT <- 5
PLOT_UNITS <- "in"
PLOT_RES <- 96 # required for png() device when units aren't "px"
```

### Load data
```{r loadData}
all2019 <- read.csv("temp_daily_2019.csv",
                    header = TRUE,
                    stringsAsFactors = FALSE)
```

## Population Statistics

### Population mean for weekly checkouts
Going up to weekly to get away from near-zero values.

```{r getWeeklyCheckouts}
# reduce data to only needed columns (because we have 5,703,704 records for 2019)
names(all2019)
cleaned2019 <- all2019[, c("ID", "CheckoutDateTime")]
rm(all2019)
head(cleaned2019)

# extract Date object from CheckoutDateTime
cleaned2019$date <- strptime(cleaned2019$CheckoutDateTime, 
                             format = "%m/%d/%Y %I:%M:%S %p")

# create new field for the week number in the year (01 - 53) 
cleaned2019$week <- strftime(cleaned2019$date, format = "%V")

head(cleaned2019) # note the POSIXlt date might not render as expected, that's ok.

# get weekly checkouts
checkoutsByWeek <- cleaned2019 %>%
  group_by(week) %>%
  summarise(weeklyCheckouts = n()) 

head(checkoutsByWeek)
```
```{r popMean_weeklyCheckouts}
print(paste0("Average weekly checkouts in 2019: ",
             round(mean(checkoutsByWeek$weeklyCheckouts))))
```
Wow! That is a lot of checkouts in a week. Let's try daily instead.

### Population mean for daily checkouts
Going down to daily checkouts because weekly is HUGE!

```{r getDailyCheckouts}
# create new field for the day
cleaned2019$day <- strftime(cleaned2019$date, format = "%m/%d/%Y")

# get daily checkouts
checkoutsByDay <- cleaned2019 %>%
  group_by(day) %>%
  summarise(dailyCheckouts = n()) 

head(checkoutsByDay)
```
```{r popMean_dailyCheckouts}
print(paste0("Average daily checkouts in 2019: ",
             round(mean(checkoutsByDay$dailyCheckouts))))
```
