---
title: "Field of Flowers - Generate Population Data"
author: "Joy-El Talbot"
date: "6/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# TODO change path to your working directory
WORKINGDIR <- "C:/Users/joyel/Documents/Roux/ALY6010/2021Sb_ALY6010/"

# R-markdown/Knitr alternative to setwd()
# MUST be run in this first chunk called "setup"
knitr::opts_knit$set(root.dir = WORKINGDIR) # sets working dir when Knitting

# TODO if coding live as you build this document copy & paste the setwd() command into your console.
setwd(WORKINGDIR) # sets working dir for R studio instance IF copied directly into console

# suppress scientific notation
options(scipen = 99)

# suppress loading strings as factors
options(stringsAsFactors = FALSE)

# set seed for randomizer
set.seed(4537)
```

## Purpose
Generate a population of dummy data for the week 3 group exercise. 

### Scenario

Imagine a field of flowers as our population. It contains flowers of different sizes and colors. Can you describe the field based on a sample of the data?

### Table of Contents
  
  1. Purpose
    a. Scenario
  2. Project Setup
  3. Data Creation
  4. Data Sampling
  
## Project Setup

### Define Constants
```{r constants}
# input source
DATADIR <- "data/"

# build results directory
# TODO name your analysis with the ANALYSISNAME constant
ANALYSISNAME <- "fieldOfFlowers"  # NO spaces!

DATESTAMP <- format(Sys.time(), "%Y-%m-%d")
RESULTSDIR <- paste0("results/",DATESTAMP,"-",ANALYSISNAME,"-results/")
dir.create(RESULTSDIR)

```
## Data Creation
Creating a field of flowers, following certain parameters (defined frequencies, means, and standard deviations). Here's an example of what the field contains.

```{r createData, echo=FALSE}
# set the core parameters for the population
flowerColors <- c("violet", "yellow", "white")
heightMeans <- c("violet" = 5, "yellow" = 8, "white" = 8)
heightSDs <- c("violet" = 3, "yellow" = 1, "white" = 3)
fractions <- c("violet" = 0.4, "yellow" = 0.4, "white" = 0.2)
fieldSize <- 10000

# function to choose ONE flower!
individual <- function(color, fraction, heightMean, heightSD){
  # pick your color
  chosenColor <- sample(x = color, 
                        size = 1, 
                        prob = fraction)
  chosenHeight <- rnorm(n = 1, 
                        mean = heightMean[chosenColor],
                        sd = heightSD[chosenColor])
  return(c(chosenColor, chosenHeight))
}

# Now let's build the population
field <- data.frame()
for (i in 1:fieldSize) {
  field <- rbind(field, 
                 individual(flowerColors, fractions, heightMeans, heightSDs))
}
names(field) <- c("Color", "Height")
field$Height <- round(as.numeric(field$Height), 2)
field[c(1:5,(fieldSize-5):fieldSize), ]
```

## Data Sampling
For each group, sample 5% of the flowers from the field.
```{r sampleData}
groups <- 4 # number of group sets to make
sampleSize <- 0.05

for (group in 1:groups) {
  rows <- sample(fieldSize, 
                 size = (sampleSize*fieldSize), 
                 replace = FALSE)
  set <- field[rows, ]
  write.csv(x = set,
            file = paste0(RESULTSDIR, "field_sample_", group, ".csv"),
            row.names = FALSE)
}
```

