---
title: "Field of Flowers - Generate Population Data"
author: "Joy-El Talbot"
date: "6/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# TODO change path to your working directory
WORKINGDIR <- "C:/Users/joyel/Documents/Roux/ALY6010/2021Sb_ALY6010/"

# R-markdown/Knitr alternative to setwd()
# MUST be run in this first chunk called "setup"
knitr::opts_knit$set(root.dir = WORKINGDIR) # sets working dir when Knitting

# TODO if coding live as you build this document copy & paste the setwd() command into your console.
setwd(WORKINGDIR) # sets working dir for R studio instance IF copied directly into console

# suppress scientific notation
options(scipen = 99)

# suppress loading strings as factors
options(stringsAsFactors = FALSE)

# set seed for randomizer
set.seed(4537)
```

## Purpose
Generate populations of dummy data for the week 4 group exercise. 

### Scenario

We suspect that a central field is the source of the spread of an invasive species of flower that happens to be purple. Unfortunately, at first glance, it looks much like a native species (also purple). We want to determine if nearby fields likely include the invasive species. 

We have a crew of eager grade school students willing to tally purple vs non-purple flowers that they see, but can’t reliably score for the invasive species beyond that. Using data from a remote field (unlikely to have been invaded) and one of the fields of concern, determine if the field is likely invaded.


### Table of Contents
  
  1. Purpose
    a. Scenario
  2. Project Setup
  3. Data Creation
  4. Data Sampling
  
## Project Setup

### Define Constants
```{r constants}
# input source
DATADIR <- "data/"

# build results directory
# TODO name your analysis with the ANALYSISNAME constant
ANALYSISNAME <- "week4_fieldsOfFlowers"  # NO spaces!

DATESTAMP <- format(Sys.time(), "%Y-%m-%d")
RESULTSDIR <- paste0("results/",DATESTAMP,"-",ANALYSISNAME,"-results/")
dir.create(RESULTSDIR)

```
## Data Creation
Creating a field of flowers, following certain parameters (defined frequencies, means, and standard deviations). Here's an example of what the field contains.

```{r createData, echo=FALSE}
# set the core parameters for the population
flowerColors <- c("purple", # native
                  "purple", # invasive
                  "yellow", "white")

remoteField <- sample(x = flowerColors,
                      size = 500, 
                      prob = c(0.3, 0, 0.4, 0.3),
                      replace = TRUE)
nwField <- sample(x = flowerColors,
                      size = 500, 
                      prob = c(0.3, 0.05, 0.4, 0.3),
                      replace = TRUE)
sField <- sample(x = flowerColors,
                      size = 500, 
                      prob = c(0.3, 0.1, 0.4, 0.3),
                      replace = TRUE)
eField <- sample(x = flowerColors,
                      size = 500, 
                      prob = c(0.3, 0.15, 0.4, 0.3),
                      replace = TRUE)

# now export the data as tables
group1 <- cbind(remoteField, nwField)
head(group1)
write.csv(group1,
          file = paste0(RESULTSDIR, "field_sample_1.csv"),
          row.names = FALSE)

group2 <- cbind(remoteField, sField)
head(group2)
write.csv(group1,
          file = paste0(RESULTSDIR, "field_sample_2.csv"),
          row.names = FALSE)

group3 <- cbind(remoteField, eField)
head(group3)
write.csv(group3,
          file = paste0(RESULTSDIR, "field_sample_3.csv"),
          row.names = FALSE)
```


