---
title: "Week 3 R Practice"
author: "Anubhav Saha"
date: "06/10/2021"
output:
  pdf_document: default
  html_document: default
---

### Introduction

Hypothesis testing is the use of statistics to determine the probability that a given hypothesis is true. The usual process of hypothesis testing consists of four steps:

1. Formulate the null hypothesis H0 (commonly, that the observations are the result of pure chance) and the alternative hypothesis Ha (commonly, that the observations show a real effect combined with a component of chance variation).

2. Identify a test statistic that can be used to assess the truth of the null hypothesis.

3. Compute the P-value, which is the probability that a test statistic at least as significant as the one observed would be obtained assuming that the null hypothesis were true. The smaller the P-value, the stronger the evidence against the null hypothesis.

4. Compare the p-value to an acceptable significance value alpha (sometimes called an alpha value). If p<=alpha, that the observed effect is statistically significant, the null hypothesis is ruled out, and the alternative hypothesis is valid.



#### Load the MASS, datasets and other required packages in R using the library function

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS) #Loading the package MASS
library(datasets)
library(dplyr)
library(ggpubr)
```


### One-sample t-test

Use the “chem” dataset to answer the question, “is the flour production company producing whole meal flour with greater than 1 part per million copper in it?”


#### code:

```{r}
chem #To glance our checm dataset
str(chem) #To get its structure
View(chem) #To view our dataset on a new window
MASS::chem #same as chem
?chem #To get help on our dataset
summary(chem) #Gives the five number summary of our dataset


hist(chem, 
     main="copper in wholemeal flour, in parts per million.", 
     xlab="Copper (ppm)", 
     border="light blue", 
     col="pink", 
     las=1, 
     breaks=5) #plots the histogram with given parameters

```

#### Step 1: Formulate Null and alternative Hypothesis

As we need to find out if there is > 1ppm of copper present, we'll use a one tailed t test

Null hypothesis (H0) : The mean value of copper present is less than or equal to one, i.e. mu<=1

Alternative hypothesis (H1) : The mean value of copper present is greater than one, i.e. mu>1 (one tailed test)


#### Step 2: Identify the test statistic

Since n < 30, we will use the t statistic instead of the z statistic.

```{r}
?t.test
```

Performs one and two sample t-tests on vectors of data.

Default arguments:
t.test(x, y = NULL,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95, ...)


#### Step 3: Compute the P-value


```{r}
ttest_chem <- t.test(chem,
                     alternative="greater",
                     mu = 1,
                     conf.level=0.99)
ttest_chem
```

```{r}
attributes(ttest_chem)
```
```{r}
ttest_chem$p.value
```
```{r}
ttest_chem$conf.int
```



#### Step 4: Compare the p-value to alpha


Since p-value < alpha (0.01), we can reject the NULL hypothesis and adopt our alternative hypothesis that mu > 1, i.e., the flour production company is producing whole meal flour with greater than 1 part per million copper in it.




### Two-sample t-test 

Use the “cats” dataset to answer the question, “do male and female cat samples have the same body weight?”

#### code:

```{r}
?cats
View(cats)
male <- subset(cats, subset=(cats$Sex=="M")) # used subset function to get separate vectors for male and female cat body weight
female <- subset(cats, subset=(cats$Sex=="F"))

ggplot(male,aes(x=Bwt)) +
    geom_histogram(aes(y=..density..),binwidth=0.25,colour="black", fill="light grey") +
        geom_density(alpha=.2, fill="red")
ggplot(female,aes(x=Bwt)) +
    geom_histogram(aes(y=..density..),binwidth=0.15,colour="black", fill="light grey") +
        geom_density(alpha=.2, fill="blue")

male_sample <- sample(male, size = 29, replace = TRUE) #Taking sample of 29 values so that n<30 to satisfy our t test constraints
female_sample <- sample(female, size = 29, replace = TRUE)

```

As we need to find out if the bwt of male cats and female cats are same or not, we'll use a two tailed t test

Null hypothesis (H0) is as follows: Male and female cat samples have the same body weight, i.e difference of mean (mud) = 0

Alternative hypothesis (H1) is as follows:  Male and female cat samples have unequal body weight, i.e difference of mean (mud) != 0



#### Two sample unpaired t test with unequal variances

```{r}
ttest_unpaired = t.test(male$Bwt, female$Bwt,
                        var.equal = FALSE,
                        paired = FALSE)
ttest_unpaired
```

**Intepretation of the results**


Since p-value < alpha (0.05), we can reject the NULL hypothesis and adopt our alternative hypothesis that mud!=0, i.e., male and female cats do not have same body weights.



###  Paired t-test 

Use the “shoes” dataset to answer the question, “did material A wear better than material B?”


#### code


```{r}
?shoes
shoes

shoe <- data.frame(shoes)
ggplot(shoe,aes(x=A)) +
    geom_histogram(aes(y=..density..),binwidth=1,colour="black", fill="light grey") +
        geom_density(alpha=.2, fill="red")

ggplot(shoe,aes(x=B)) +
    geom_histogram(aes(y=..density..),binwidth=1,colour="black", fill="light grey") +
        geom_density(alpha=.2, fill="blue")

```


As we need to find out if the material A wears better than material B, we'll use a one - tailed two sampled paired t test

Null hypothesis (H0) is as follows: Shoe of materal A wears worse or equal to the shoe of material B 
Alternative hypothesis (H1) is as follows:  Shoe of materal A wears better than shoe of material B.


#### two sample paired t test

```{r}
paired = t.test(shoes$A, shoes$B,
                paired=TRUE,
                alternative="greater")
paired
```

**Intepretation of the results**

Since p-value > alpha (0.05), we fail to reject the NULL hypothesis i.e.at 95% confidence level, there is not enough evidence to deny that Shoe of materal A wears worse or equal to the shoe of material B 



### Independent two sample t-test

Link to the dataset lung capacity https://drive.google.com/file/d/0BxQfpNgXuWoIWUdZV1ZTc2ZscnM/edit?resourcekey=0-gqXT7Re2eUS2JGt_w1y4vA


```{r}
Data <- read.csv(file.choose())
View(Data)
```

```{r}
names(Data)
```

```{r}
boxplot(Data$LungCap ~ Data$Smoke)
```

Ho : Mean Lung capacity of smokers = that of non smokers i.e. difference in mean = 0
Ha : Mean Lung capacity of smokers != that of non smokers i.e. difference in mean != 0

Based on the boxplot, non smokers seem to have a higher variance.

Let's confirm that by getting the variance


```{r}
var(Data$LungCap[Data$Smoke=="yes"])
var(Data$LungCap[Data$Smoke=="no"])
```


#### Two sided t test

```{r}
t.test(Data$LungCap ~ Data$Smoke)
```

If we wanna test for difference in mean more than 0, we can provide that argument during the t test



